\documentclass[12pt]{article}
\usepackage[margin=0.75in]{geometry}
\usepackage{fontspec}
\setmainfont{Latin Modern Roman}
\usepackage{microtype}
\usepackage{multicol}
\usepackage{fancyhdr}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\setlength{\columnsep}{0.28in}
\setlength{\parindent}{1.1em}
\setlength{\parskip}{0.35em}
\setlength{\headheight}{15pt}
\titleformat{\section}{\large\bfseries\scshape}{}{0pt}{}
\titleformat{\subsection}{\normalsize\bfseries}{}{0pt}{}
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0.4pt}
\newcommand{\focusbox}[1]{\par\vspace{0.4em}\noindent\begingroup\setlength{\fboxsep}{6pt}\colorbox{gray!12}{\begin{minipage}{\dimexpr\linewidth-2\fboxsep\relax}#1\end{minipage}}\endgroup\par\vspace{0.5em}}
\newcommand{\studentline}{\vspace{0.35em}\hrule\vspace{0.65em}}

\fancyhead[L]{Whately Elements of Logic}
\fancyhead[R]{Teacher Guide}
\begin{document}
\begin{center}
{\LARGE\bfseries Teacher Guide: Week 2}\par\vspace{0.25em}
{\large\scshape One Word, Two Ideas: Why Ambiguity Breaks Arguments}\par\vspace{0.4em}
{12th Grade Long Logic Unit Based on Richard Whately's \textit{Elements of Logic}}\par\vspace{0.6em}
\rule{0.82\textwidth}{0.4pt}
\end{center}

\section*{Week 2 Overview}
Week 2 builds directly on the four-step analysis habit from Week 1 by focusing on one of the most common reasons arguments fail the “test” step: ambiguous or shifting terms. Students learn to notice when a key word is being used in two different senses (equivocation) and to fix the meaning of important terms before evaluating an argument. The week draws from Whately's treatment of terms and ambiguity while remaining practical and example-driven. In the revised twelfth-grade plan, students begin a recurring habit of testing words from \textit{Macbeth} and \textit{Julius Caesar}: fate, manhood, honor, ambition, liberty, and tyranny.

\section*{Essential Question}
How can the same word mean two different things inside a single argument, and what can we do about it?

\section*{Student-Facing Materials}
\begin{itemize}[leftmargin=*]
\item \textit{One Word, Two Ideas: Why Ambiguity Breaks Arguments} — student reading handout.
\item \textit{Week 2 Argument Lab: Fixing the Term} — practice identifying ambiguous terms, repairing arguments, and creating original examples.
\item \textit{Macbeth Lit-Anchor: Words That Lie Like Truth} — Shakespeare anchor passage for testing Macbeth's confidence in ambiguous prophecies: ``none of woman born'' and ``Birnam wood ... shall come.''
\end{itemize}

\section*{Learning Goals}
Students will be able to:
\begin{enumerate}[leftmargin=*]
\item Define a term and explain how the same word can refer to different ideas.
\item Distinguish ambiguity from equivocation.
\item Identify when an argument relies on an unstated shift in the meaning of a key term.
\item Fix the meaning of a disputed term before testing an argument.
\item Repair an equivocal argument by making the term precise.
\item Recognize that some apparent disagreements are actually disagreements about the meaning of words.
\item Build a portfolio term investigation using either a contemporary argument or a recurring Shakespeare anchor.
\end{enumerate}

\section*{Key Vocabulary}
term, meaning, ambiguity, equivocation, fix the term, context

\section*{Weekly Lesson Sequence}

\subsection*{Day 1: Opening Question and First Read}
\begin{itemize}[leftmargin=*]
\item Begin with: “Can two people use the exact same word and mean completely different things?” Have students write a quick example.
\item Introduce the idea that many arguments look strong only because a word quietly changes meaning.
\item Read the first half of the student reading aloud or in pairs.
\item Stop after the “bank” example. Ask: Why does the argument sound silly once the two meanings are pointed out?
\end{itemize}

\subsection*{Day 2: Finish Reading and Annotate Examples}
\begin{itemize}[leftmargin=*]
\item Finish the reading.
\item Students mark every example and label the two different meanings of the key term.
\item Complete Reading Focus questions 1–4.
\end{itemize}

\subsection*{Day 3: Term-Fixing Practice}
\begin{itemize}[leftmargin=*]
\item Use the Argument Lab. Students work in pairs to identify the ambiguous term in each argument and propose a fixed definition.
\item Discuss: How does fixing the term change whether the conclusion follows?
\end{itemize}

\subsection*{Day 4: Macbeth Lit-Anchor: Words That Lie Like Truth}
\begin{itemize}[leftmargin=*]
\item Read the guided Act 4, Scene 1 apparitions passage from the lit-anchor.
\item Put two phrases on the board: ``none of woman born'' and ``Birnam wood ... shall come.'' For each, ask: What meaning does Macbeth assume? What alternate meaning remains possible?
\item Emphasize that Macbeth's confidence depends less on a simple falsehood than on an unfixed phrase taken in the most comforting possible sense.
\item Students begin the \textit{Macbeth Equivocation Chart} portfolio artifact.
\end{itemize}

\subsection*{Day 5: Discussion, Repair, and Exit Task}
\begin{itemize}[leftmargin=*]
\item Read the independent Act 5, Scene 8 excerpt if it was not finished on Day 4: Macduff's revelation and Macbeth's ``palter with us in a double sense'' response.
\item Class discussion: “When have you seen real arguments (in the news, social media, school, or literature) that depend on ambiguous terms?”
\item Students finish either one lab repair or the \textit{Macbeth Equivocation Chart}.
\item Exit task: Students write a short portfolio reflection explaining why fixing terms matters for the four-step analysis they learned in Week 1, naming one Shakespeare term that requires careful definition.
\end{itemize}

\section*{Lit-Anchor Teaching Notes}
The Week 2 lit-anchor gives students the central literary version of equivocation. Macbeth is not merely tricked by statements that are false on their face. He is ruined by phrases that keep one promise to his ear while allowing another meaning in reality. That makes the passage especially useful for Whately's lesson: fix the terms before trusting the conclusion.

\begin{itemize}[leftmargin=*]
\item \textbf{Guided passage:} \textit{Macbeth} Act 4, Scene 1, the apparitions' warnings and promises.
\item \textbf{Independent passage:} Act 5, Scene 8, Macbeth and Macduff, ending with Macbeth's recognition that the fiends ``palter with us in a double sense.''
\item \textbf{Central logical pressure:} Macbeth hears broad protection where the words allow narrow exceptions.
\item \textbf{Portfolio artifact:} Macbeth equivocation chart: phrase, Macbeth's interpretation, alternate meaning, hidden assumption, effect on reasoning, and connection to Whately.
\end{itemize}

For ``none of woman born,'' students should see Macbeth's assumed meaning: no human being can harm him. Macduff exposes the alternate meaning: not born in the ordinary manner. For ``Birnam wood ... shall come,'' Macbeth assumes a whole forest must uproot itself; the words can also be fulfilled by soldiers carrying branches. The common hidden assumption is that Macbeth's first comfortable meaning is the only possible meaning.

\section*{Notes on Examples}
The student reading deliberately uses a mix of light, obvious examples (“bank”) and more serious, debatable examples (“liberal education,” equality, and tolerance). The lab adds \textit{Macbeth} and \textit{Julius Caesar} examples so students see that ambiguity is not only a trick in silly arguments but a real issue in literature, public speech, and substantive disagreements. Encourage students to bring their own examples from current events, school debates, or Shakespeare.

\section*{Connection to Future Weeks}
This week prepares students for later work on propositions, syllogisms, fallacies (many of which rely on equivocation), and verbal versus real disputes. The habit of fixing terms will also make the study of propositions and syllogisms clearer when those weeks arrive.
\end{document}